---
title: "STAT 380 Final Proj"
format: html
editor: visual
---

## Introduction

The Marvel Universe is one of the largest fictional communities in history. With thousands of characters created over the last 80+ years, "Appearances" serve as a key for seeing a character's success, staying power, and cultural relevance. This data is important to us as a group because it allows us to look at the stories of characters and what that may mean for their longevity in the Marvel Universe. By looking at which traits relate to a higher number of appearances, we can see what types of characters are more popular with the Marvel audience over others.

## Data Provenance - 5 W's

#### Who

The data was originally found on the site Marvel Wiki, an online database ran by fans of Marvel. 

#### What

It is a csv dataset that contains more than 16,000 characters with attributes that include their identity status, gender, year they first appeared, etc.

#### Where

The source is Marvel Wiki, a Marvel database that is a fan edited "encyclopedia" that documents Marvel books and Cinematic Universe (MCU) media.

#### When

The data includes characters from 1939 all the way to 2014, which is when the data was pulled.

#### Why

The data was originally collected and used to look at character diversity and the evolution of the comic company.

#### How

Online tools were used to pull the data about each character that is included in the dataset.

#### Notes:

Because the dataset was maintained by fans, some of the characters have missing data and their appearances count is based off of the number of individual book issues that a character was tagged in on the origin website.

## EDA

```{r, echo=FALSE}
# 1. Load libraries and dataset
library(tidyverse)

marvel_data <- read_csv("marvel-wikia-data.csv")

# 2. Data Cleaning
marvel_clean <- marvel_data %>%
  filter(!is.na(APPEARANCES), !is.na(Year)) %>%
  mutate(
    ALIGN = ifelse(is.na(ALIGN), "Unknown", ALIGN),
    ID = ifelse(is.na(ID), "Unknown", ID),
    SEX = ifelse(is.na(SEX), "Unknown", SEX)
  ) %>%

  mutate(across(c(ALIGN, ID, SEX), as.factor))

# 3. EDA

# Plot 1: Relationship between Year and Appearances (Scatter Plot)
ggplot(marvel_clean, aes(x = Year, y = APPEARANCES)) +
  geom_point(alpha = 0.2, color = "steelblue") +
  geom_smooth(method = "lm", color = "red") +
  theme_bw() +
  labs(
    title = "Character Appearances vs. Year of Introduction",
    subtitle = "Older characters tend to have significantly more appearances",
    x = "Year",
    y = "Number of Appearances"
  )

# Plot 2: Appearances 
ggplot(marvel_clean, aes(x = ALIGN, y = APPEARANCES, fill = ALIGN)) +
  geom_boxplot() +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Distribution of Appearances by Character Alignment",
    x = "Alignment",
    y = "Appearances (Log Scale)"
  )

# Plot 3: Summary Table
marvel_summary <- marvel_clean %>%
  group_by(ALIGN) %>%
  summarise(
    Count = n(),
    Average_Appearances = mean(APPEARANCES),
    Median_Appearances = median(APPEARANCES)
  )

print(marvel_summary)
```

The exploratory analysis of the Marvel dataset shows an interesting distribution for the target variable of appearances. While a large majority of characters have less than ten recorded appearances, a small elite group of super characters like Spider-Man and Captain America, feature counts in the thousands.

A scatter plot of appearances with the year included gives a strong suggestion that characters introduced in the mid-20th century have naturally accumulated more exposure over time because of how long they have been around. The box plots also show that character traits such as Alignment and Identity Status may influence these counts.

For example, "Good" characters with "Secret Identities" appear to have a higher average appearance vs the "Bad" or "Public" ones. This suggests that while time heavily influences character longevity, personal attributes also play a significant role. This justifies the use of both simple linear regression to capture the temporal trend and a regression tree to uncover the more complex influences of character traits.


## Regression Based Algorithm: Multiple Linear Regression Model

The first model created to be able to see which attributes are effective in predicting the amount of appearances for a character, will be created by using Multiple Linear Regression. The process of Multiple Linear Regression is finding a model that relates the response variable, appearances of a character, to multiple different predictors linearly. Those predictors will be chosen in the process below. 

The first regression model created from the model selection uses the attributes: `ALIGN` (Alignment of the character), `EYE` (Eye color), `ALIVE` (If the hero is currently alive or not), `Year` (Year of which the character was introduced).

Before creating this model, some variables were deliberately removed from the model selection. The variables removed were `page_id`, `name` and `urlslug` for all being variables for identification and not actually about the characters themselves. As well as `GSM`, which gave the sexual identity of the character. This was remove due having primarily NA values for almost all the observations, and otherwise would not prove useful to include because of it. The observations that had NA values for the predictors that could end up in the model were also removed to be able to do the variable selection properly.

After the first model, before confirming that these were the variables to be used to predict `APPEARANCES`, one of the variables had a lot of insignificant coefficients. The variable was `EYE`, which is the eye-color of the character. Due to having 23 levels within the variable, keeping the variable in the model would ultimately be prone to over fitting and also just adds too much complexity to the model.

After all of the analysis and reconstruction of the model, the final model included: `ALIGN`, `ALIVE`, and `Year` as the most significant attributes in predicting the appearances of a character. The root mean square error, which describe how much the mean difference between the predicted amount of appearances and the actual appearance values, is 95.96. That means that we are off on average by 96 times less or more than a character actually showed up. Which is not a terrible margin of error.

```{r, echo = FALSE, results = 'hide'}
library(tidyverse)
library(caret)

#load data
marvel_data <- read.csv("marvel-wikia-data.csv")

#Cross Validation
  set.seed(161) # setting seed to be able to reproduce our code
  
  
#Cross Validating our dataset
  cv <- trainControl(
    method = "cv",
    number = 10
  )

#base model, created from what data exploration showed to be most 
model1_CV <- train(
  APPEARANCES ~ Year,
  data = marvel_data,
  method = 'lm',
  trControl = cv,
  na.action = na.omit
)

#Checking the coefficients for the first model
model1_CV$finalModel$coefficients

#grabbing the rmse value
model1_CV$results$RMSE

#Model Selection using Backward Selection using AIC

#removing al overvations with na values for the possible predictors
marvel_data <- na.omit(marvel_data[,c("APPEARANCES","ID", "ALIGN", "EYE", "HAIR", "SEX", "ALIVE", "Year")])

#removing page_id, name and urlslug due to irrelevancy, removing GSM due to many NA values
full_marvel_model <- lm(APPEARANCES ~ ID+ALIGN+EYE+HAIR+SEX+ALIVE+Year, data = marvel_data)

stats::step(
  object = full_marvel_model,
  data = marvel_data,
  direction = "backward",
  k = log(nrow(marvel_data)), #BIC parameter
  trace = 1
)

#Model 2:
model2 <- lm(APPEARANCES ~ ALIGN + EYE + ALIVE + Year, data = marvel_data)

#looking at the coefficients
#summary(model2)
rmse2 <- sqrt(mean(model2$residuals^2))

#checking the value
#rmse2

#removing EYE <- due to multiple insignificant predictors
model3 <- lm(APPEARANCES ~ ALIGN + ALIVE + Year, data = marvel_data)
#summary(model3)
rmse3 <- sqrt(mean(model3$residuals^2))
```

#### Strength and Weaknesses of the Regression

The main weakness of using a Multiple Linear Regression model for this data in particular is that majority of the attributes in the data set are categorical. So, trying to fit a linear relationship with so little quantitative variables makes selecting variables tough, as well as make our model less accurate because categorical variables don't have a specified quantity that is affecting the slope. This could be fixed by finding additional data with more quantitative variables such as height, or weight to create the model, however we were unable to find any suppliment data for our original data set. Strengths of the Multiple Linear Regression model, are it is a much simpler model to understand and explain. Variable selection is relatively simple and can easily be explained to someone who doesn't have much knowledge on the subject. 